|

### 2021-11-18 07-Chp-Triangulation.Rmd
End of day:

@@

<!----------------------------------------------------------------------------->
(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases.

@@

```{r exampleDirection, include=F}
```

```{r exampleDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:exampleDirection-cap)', out.width='100%', fig.scap='(ref:exampleDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/exampleDirection.png"))
```
<!----------------------------------------------------------------------------->

@@

Following definition of the extent, type and direction of bias in each domain, I assigned a prior distribution on the log scale to each combination. For example, a "high" additive bias in $j$th domain of the $i$th study is deifined as $\delta_{i,j}^{\mathrm{High}}$ and uncertainty around this estimate is given by a distribution:

@@

The sign of $\mu_{i,j}^{\mathrm{High}}$ is defined by the absolute direction of the bias. If the bias is expected to pull the effect to the left, the position is given a negative sign, and if it is expected to pull the effect to the right, it is given a positive sign.

One key feature of the domain-based risk-of-bias domains is that the domains are considered interchangeable - i.e. a "high" judgement in one domain is equivalent to a "high" judgement in any other. As such, for this analysis, I assumed that the distributions assigned to each level of bias are identical across all values of $j$, that is, are identical for all domains in the risk of bias assessment tool. To illustrate, for additive biases these means:

@@

<!-- TODO In the absence of any empirical data regarding the priors, I used the data from the expert elicication approach to calculate both the mean and max. These estimates, in addition to conversations with topic experts -->

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the adjustment distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained. In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate, but does increase uncertainty around it.

@@

An identical approach was employed to assess and assign prior distributions to the risk of indirectness in each result. The target question for each result was compared against the causal question of interest with respect to three domains (population, intervention/exposure, outcome), <!-- TODO CITATION NEEDED --> again using the scale of "Low/"Moderate"/"High" to quantify the extent of indirectness.

A similar approach was used to define prior distributions for external bias or indirectness. I used a similar enforced scale of "Low/"Moderate"/"High" indirectness versus each of the 4 aspects of the target causal question. (recap). All indirectness were defined _a priori_ as being proportional in nature, in line with previous work on this topic.[@turner2009]

@@

As per the method reported in, total additive and proportional bias and indirectness were calculated for each result. Briefly,

@@

(ref:biasDirectionSingle-scap) Bias direction plot for a single study
```{r biasDirectionSingle, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:biasDirectionSingle-cap)', out.width='100%', fig.scap='(ref:biasDirectionSingle-scap)'}
knitr::include_graphics(file.path("figures/misc/UoBcrest.pdf"))
```
<!----------------------------------------------------------------------------->

@@

```{r ldlAdBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlAdBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlAdBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD.png"))
```
<!-- TODO Bias direction plot is not producing subgroup test for some reason -->
&nbsp;<!----------------------------------------------------------------------->

The results of the unadjusted and the bias-/indirectness-adjusted (under the Scenario 1 of additive biases) are shown in Figure \@ref(fig:fpLdlAd). I observed no substantial change to the overall effect (unadjusted `r unadj_effect`; adjusted `r adj_effect`), though adjustment for bias and indirectness substantially reduced the heterogeneity between studies (unadjusted $\tau^2$ = `r unadj_tau2`, $I^2$ = `r unadj_I2`; adjusted  $\tau^2$ = `r adj_tau2`, $I^2$ = `r adj_I2`). However, as can been seen from the right panel in the figure, this reduction in heterogeneity between studies is largely achieved via a reduction in precision.

@@

@@

@@

@@

Additional if users are interested in only a summary of the evidence rather than assessing results in relation to a target qusetion, then internal biases can be adjusted for on their own.

In previous attempts at bias/indirectness-adjusted meta-analysis, the extent to each was assessed via a elicitation process using a number of experts. This approach is potentially subject to misclassification of the impact of bias on the basis of the results, as there is no way to ensure that results at a similar level of bias for confounding (for example) are being adjusted by the same amount. For example, if experts are influenced by the absolute effect estimate, then a result demonstrating a stronger protective effect may receive greater modification than a study at the same "risk" of confounding but with a more modest effect. This is particular true where experts have preconceived ideas about the true effectiveness of the intervention.

In this case, separating the assessment of bias from the assignment of modifying values to each level of bias will limit the potential for. Alternatively, reasonable modifying distributions for each level of bias could be defined _a priori_ by the study team, similar to how important confounders and co-interventions are defined in advance when performing ROBINS-I/ROBINS-E assessments.

@@

@@

I hope this presents step forward in how researchers think about and visualise triangulation at the result level, rather than simply saying that certain

@@

The implications of th synthesised work of this thesis presened in this chapter to clinical practice, public health and future research is considered in the following chapter.

>>>>

<!-- TODO Talk about how the CPRD and IPD analysis results were assessed here! -->

<!-- Risk of bias assessment was performed for each of the included cohorts using the risk-of-bias assessment tool for non-randomised studies of exposures introduced previously (for a more detailed discussion of this tool, see Section \@ref(rob-tools-nrse)). -->

<!-- Due to the inherent conflict of interest in assessing my own analysis, the result were also assessed by two independent reviewers. -->

@@

<!----------------------------------------------------------------------------->
 ```{r exampleDirectionSetup, include = F}
 ```

(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases.

@@

```{r exampleDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:exampleDirection-cap)', out.width='100%', fig.scap='(ref:exampleDirection-scap)'}

knitr::include_graphics(file.path("figures/tri/exampleDirection.png"))

```
<!----------------------------------------------------------------------------->

@@

Following definition of the extent, type and direction of bias in each domain, I assigned a prior distribution on the log scale to each combination. For example, a "high" additive bias in $j$th domain of the $i$th study is defined as $\delta_{i,j}^{\mathrm{High}}$ and uncertainty around this estimate is given by a distribution:

@@

The sign of $\mu_{i,j}^{\mathrm{High}}$ is defined by the absolute direction of the bias. If the bias is expected to pull the effect to the left, $\mu_{i,j}^{\mathrm{High}}$ is given a negative sign, and if it is expected to pull the effect to the right, it is given a positive sign.

One key feature of the domain-based risk-of-bias tools is that the domains are considered interchangeable - i.e. a "high" judgement in one domain is equivalent to a "high" judgement in any other. As such, for this analysis, I assumed that the distributions assigned to each extent of bias were identical across all values of $j$, that is, were identical for all domains in the risk-of-bias assessment tool. To illustrate for additive biases:

@@

<!-- TODO In the absence of any empirical data regarding the priors, I used the data from the expert elicitation approach to calculate both the mean and max. These estimates, in addition to conversations with topic experts -->

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the adjustment distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained (e.g. for an unpredictable "high" assessment, the distribution would be $N(0,0.12)$). In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate, but does increase the uncertainty around it.

@@

An identical approach was employed to assess and assign prior distributions to the risk of indirectness in each result. The target question for each result was compared against the causal question of interest with respect to three domains (population, intervention/exposure, outcome), <!-- TODO CITATION NEEDED --> again using the scale of "Low/"Moderate"/"High" to quantify the extent of indirectness. All indirectness were defined _a priori_ as being proportional in nature, in line with previous work on this topic.[@turner2009]

@@

Using the the method reported in Turner _et al._,[@turner2009] total additive and proportional bias and indirectness were calculated for each result. Briefly,

@@

(ref:biasDirectionSingle-scap) Bias direction plot for a single study
```{r biasDirectionSingle, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:biasDirectionSingle-cap)', out.width='100%', fig.scap='(ref:biasDirectionSingle-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD_single.png"))
```
<!----------------------------------------------------------------------------->

@@

```{r ldlAdBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlAdBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlAdBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD.png"))
```
<!-- TODO Bias direction plot is not producing subgroup test for some reason -->

The results of the unadjusted and the bias-/indirectness-adjusted (under the Scenario 1 of additive biases) are shown in Figure \@ref(fig:fpLdlAd). I observed no substantial change to the overall effect (unadjusted `r unadj_effect`; adjusted `r adj_effect_scenario1`), though adjustment for bias and indirectness substantially reduced the heterogeneity between studies (unadjusted $\tau^2$ = `r unadj_tau2`, $I^2$ = `r unadj_I2`; adjusted  $\tau^2$ = `r adj_tau2`, $I^2$ = `r adj_I2`). However, as can been seen from the right panel in the figure, this reduction in heterogeneity between studies is largely achieved via a reduction in precision.

@@

While this analysis may be biased from the non-empirical bias adjustment factors, at the very least, it acknowledges the the uncertainty of evidence via the widening of confidence intervals. 
No less problematic than the analysis fo biased estimates in a meta-analysis which does not account for bias.

@@

Allowing for the strong assumption that

@@

#### Subjectivity of bias-/indirectness-assessment

<!-- TODO Add text here on the subjectivity of qualitative risk of bias assessment and cite ROBINS-I to say that risk of bias is hard. Note that this will apply  -->

A key issue in this analysis is t

As observed from Figure \@ref(@) in the bias-adjusted analysis, the Ostergaard _et al._ 2017

@@

Additional if users are interested in only a summary of the evidence rather than assessing results in relation to a target question, then internal biases can be adjusted for on their own by setting the indirectness distributions to $N(0,0)$.

The extent of a bias in a domain is unlikely to map accurately across the same domains in different studies, which

It for this reason that in previous attempts at bias-/indirectness-adjusted meta-analysis, the extent of bias in each study was assessed via a elicitation process using a panel of methodologists and topic experts.[@thompson2011; @turner2009; @wilks2011] However, this approach is potentially subject to misclassification of the impact of bias on the basis of the results, as there is no way to ensure that results at a similar level of bias for confounding (for example) are being adjusted by the same amount across studies. As an illustrative example, if experts are influenced by the knowledge of the effect estimate in a study, then a result demonstrating a stronger protective effect may receive greater modification than a study at the same "risk" of confounding but with a more modest effect. This is likely to be part where experts have preconceived ideas about the true effectiveness of the intervention.

In this case, separating the assessment of bias from the assignment of modifying values to each level of bias will limit the potential for. Alternatively, reasonable modifying distributions for each level of bias could be defined _a priori_ by the study team using the elicitation procedure detailed previously,[@turner2009] similar to how important confounders and co-interventions are defined in advance when performing ROBINS-I/ROBINS-E assessments.[@sterne2016]

@@

Future research could and should also take into account bias at the analysis level in terms of missing evidence. For example, the set of previously published NRSE included in the case study were found to be at high risk of bias from missing evidence (see Figure \@ref(fig:lipidFractionsAD) in Chapter \@ref(sys-rev-results-heading)). Ideally, a quantitative triangulation framework would further account for this proportional evidence-level bias (as bias due to missing evidence is most likely to bias away from the null due to publication bias mechanisms), in addition to result-specific biases.

<!-- QUESTION - Can you meta-analyse meta-analysis results? As in, have a two-stage approach where you pool by evidence source, investigate bias at analysis level within evidence source, adjust these within-evidence-source results for ROB_ME bias, then pool them for final estimate? -->

@@

I hope this presents step forward in how researchers think about and visualise triangulation at the result level, rather than simply saying that certain evidence sources a

@@

The implications of the synthesised work of this thesis presented in this chapter to clinical practice, public health and future research is considered in the following chapter.
